Numerical trajectory patterns of floating macroalgae

### 2021-01-09 thesis_chapter_3.Rmd
updated discussion added. more to come

>>>>

The treatment of drifting macroalgae as purely lagrangian will not be able to accurately determine patterns of passive dispersal, such as entrainment in eddies or vortices. Although past research has shown the inclusion of inertia and windage greatly increase observed macroalgae trajectory and entrainment characteristics, however current research has focused on large macroalgae-rafts that can vary greatly in size and surface area. None of the past research has taken advantage of the advancement in numerical ocean models and lagrangian trajectory modeling to investigate how aspects of drag (hydrodynamic and wind) affect the trajectory patterns of solitary drifting macroalgae. This study compared particles trajectories with simulations that included no forms of drag with simulations that included various forms of both hydrodynamic and wind drag into account. The passive simulations deviated from passive particles trajectories when including varing levels of hydrodynamic and wind drag scenarios into account. The results from this study show the inclusion of hydrodynamic and wind drag improves entrainment patterns and as well our understanding of how surface area plays a role in solitary drifting macroalgae trajectory, in this case *E.maxima*.

The general flow in the study domain is north-westward and is characterised by sub-mesoscale processes such as vortices and sub-mesoscale eddies.

Past research has inferred from direct and indirect techniques that drifting macroalgae tend to follow general flow patterns in the region in question. For example...

The results from this numerical experiment confirm that drifting macroalgae do follow the surface current in the study region when considering only hydrodynamic drag. In addition, not only are the trajectories similar but so are the end points at the end of simulation. However, when ad-hoc wind drag (windage) is considered in combination with hydrodynamic drag the trajectories differ greatly when compared to purely langrangian particle trajectories. The inclusion of wind drag causes the particles to flow further away from the coast and with a greater distance travelled from the release location. The difference in trajectory and distance covered can be attributed to the inclusion of wind drag. Wind drag adds to the effects of hydrodynamic drag when the wind direction and current direction are opposite to each other. However, when the wind direction is in the same or similar direction to that of the current, momentum energy from the wind is transferred to the object. This causes the particles to travel further and to be exposed to different time varying flows. insert putman study here. This is also reflected in the analysis comparing distances travelled by all the particles in each simulation. The particles in the simulations that included wind drag travlled significantly greater distances compared to the passive or hydrodynamic only drag.

The addition of drag forces into the simulation also causes the particles to cluster together along the mean trajectory compared to that of the passive particles which are more evenly dispersed. This is most likely due to the shared drag characteristics among the particles which included the plants weight and surface area into the calculation. Other studies investigating flotsam trajectory characteristics have found similar results. A study by @miron2020 conducted field experiments using a range of objects such as spheres, cubes, cuboids to compare to investigate the effect of inertia on particles dynamics. The results from the study showed that objects tended to cluster according to the shape of the object which the authors attributed to shared characteristics of inertia. Although drag forces were used in this study and not inertia, this study provides evidence to account for the higher density of particles along the mean trajectory.

The ocean is made up of varying time-varying flows and changes in an objects velocity can cause the object to be exposed to different flow patterns over time. The lack of dissimilarity between different cross-sectional area types (minimum, mean and maximum) was an unexpected result. A similar result was found by @gouvello2020 who investigated the effects of swimming behavior on sea turtle hatchling dispersal in the Agulhas region. The results suggested that the hatchlings trajectories are mostly influenced by the ocean currents in the first year of hatching due to low swimming speeds of juveniles. Therefore, the differences in cross-sectional area types may not be significant enough in order to have a significant effect on the relevant velocity vectors (zonal and meridonal). Another possible reason is the resolution of the ocean model used. The resolution of the underlying ocean model is an important aspect of lagrangian ocean modeling and must be able to resolve sub-grid scale ocean processes for finer scale applications. A study by @hart2018 assessed the inclusion of stochastic motion, wind and currents into forecasting for search and rescue using the same model ocean model used in this study. The authors found that the inclusion of brownian motion greatly increased the accuracy in representing sub-grid scale processes and that the inclusion of wind, currents and stochastic motion greatly improved forecasting applications. Therefore, the authors argue that the lack of dissimilarity between simulations of different cross-sectional area and drag exposure scenarios is not as a result of the resolution of the model.

The same is also true when comparing different hydrodynamic and wind drag exposure scenarios. These scenarios reflect different buoyancy situations which result in different amounts of surface area exposed to hydrodynamic and/or wind drag. The similarities between the different combined drag scenarios suggests that higher magnitudes of wind exposure do not significantly alter trajectories. Instead, the results from this study suggest that the inclusion of drag forces in simulating macroalgal trajectory may result in improving accuracy of entrainment patterns. All the simulations that included any form of drag resulted in a fairly high density of particles entrained in a vortex. The entrainment and expulsion from eddies and vortices is an important characteristic of floating macroalgae. Accuracy in predicting entrainment and expulsion of macroalgae-rafts from eddies has been investigated previously by @putman2018 and @putman2020, who show that including windage greatly improves entrainment patterns when comparing virtual and real sargassum rafts. However, the aforementioned studies investigated large rafts and not floating individuals. Inertia is based on the size/weight of the raft, while hydrodynamic and wind drag are based on current velocity and surface area. Therefore, the role of inertia may be negligible for a solitary individual and rather other forms of drag should be considered. The results from this study suggest that the inclusion of hydrodynamic and/or wind drag is an important component when simulating virtual kelp particles will greatly increase the accuracy of entrainment and expulsion from sub-grid mesoscale oceanographic features, such as eddies and vortices.

??the study by putman which used different ad-hoc windage paramterisations showed that higher windsage resulted in greater accuracy of entrainment in eddies??.
